\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2013-09-06 paper.tex
Updated style of the 3rd HATP plan
Merged Rachid's comments

Human-robot interaction is a challenge for artificial intelligence. This field
indeed lays at the crossroad of several other domains of AI and requires to
tackle them in a holistic manner: Modeling humans and human cognition;
acquiring, representing, manipulating in a tractable way abstract knowledge;
reasoning on this knowledge to make decisions; and eventually instantiating
those decisions into physical actions in coordination with humans. Many AI
techniques are invited, from visual processing to symbolic reasoning, from task
planing to \emph{theory of mind} building, from reactive control to action
recognition and learning.

We do not tackle all these issues. This article attempts however to organize
them into a coherent challenge for artificial intelligence, and also explicit
some of the paths that have been investigated on our robots and that result in a
deliberative, knowledge-oriented, architecture for human-robot interaction.

We scope our work to the robot which is
expected to achieve interactive object manipulation, fetch and carry tasks and
similar tasks in a domestic environment (Figure~\ref{fig:hri-dec}).

Two agents (a human and a robot) share a common space and exchange information
through multiple modalities. The robot is expected to explicitly take into
account the human and to manage its interactions with him.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.9\columnwidth]{grounding_robot.pdf}
\caption{Robot reasoning about HRI and anticipation of human activities.
  Sources of information are multi-modal dialogue, and monitoring of the
  environment and human activity. The robot must adapt on-line its behaviours
  by merging computed plans with reactive control.}
\label{fig:hri-dec}
\end{figure}

@@

The remaining of the article discuss the robotic architecture we have built to
tackle the AI challenges of human-robot interaction, and how its relates to
other approaches. We propose to organize this discussion in four sections.

The next section presents the overall architecture and the knowledge model we
have developed for our robots. Section~\ref{sec:impl} gives some insights on
the each of the main components of the architecture, and try to highlight their
significance for artificial intelligence. Section~\ref{sec:expe} presents two
experiments that illustrate in a concrete way what can be currently achieved
with our robots.  Section~\ref{sect|conclusion} finally summarizes our main
contributions and relate them to other recent researches in the field.

@@

\begin{figure*}
        \centering
        \includegraphics[width=0.8\textwidth]{archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

@@

This activity is however not the backbone of our architecture. Other
activities are conducted in parallel, without being explicitly considered as
desires: assessment of the situation and the environment, dialogue (including
performative dialogue that can possibly change the internal state of the robot,
but does not lead to the creation of desires, like statement assertion or
question answering), various background monitoring and recognition tasks, etc.

@@

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{spark-oro.png}
    \caption{Functional overview of knowledge base ({\sc Oro} server, top part)
    and the geometric situation assessment module (\emph{{\sc Spark}}, bottom
    part). {\sc Spark} computes symbolic relationships between objects and
    agents, and feeds them to the knowledge base.}

@@

   \caption{% In this situation, there are three tapes on the table. Two tapes
       are only reachable by the robot: the \concept{LOTR\_TAPE} (black in the 3D
       model) and \concept{GREY\_TAPE}. The third tape \concept{WALLE\_TAPE} (white
       in the 3D model) and the trashbin \concept{PINK\_TRASHBIN} are only
       reachable by the human \concept{HUMAN1}. All tapes are on the table
   \concept{TABLE}. The corresponding symbolic facts are listed in Table~\ref{table|beliefsfig7}.  }%

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\columnwidth]{second_plan.pdf}
  \caption{A plan with the wasted time social rule}
  \label{plan_hatp2}
\end{figure}

@@

\begin{figure*}[thpb]
  \centering
  \includegraphics[width=1.0\textwidth]{plan3.pdf} \\
  \caption {A plan produced by HATP to execute the high-level order ``clean the table''. Two streams of actions are generated: for the human (top) and for the robot(bottom). Synchronization points ensure the coordination.}
  \label{plan-etat2}
\end{figure*}

@@

Numerous approaches exist, like amodal \emph{proxies}~\cite{Jacobsson2008},
grounded amodal representations~\cite{Mavridis2006}, semantic maps
(\cite{Nuechter2008, Galindo2008,Blodow2011}) or
affordance-based planing and object classification~\cite{Lorken2008,
Varadarajan2011}.

@@

@@

>>>>

Human-robot interaction is a challenge for artificial intelligence. This field
indeed lays at the crossroad of several other domains of AI and requires to
tackle them in a holistic manner: Modeling humans and human cognition;
acquiring, representing, manipulating in a tractable way abstract knowledge;
reasoning on this knowledge to make decisions; and eventually instantiating
those decisions into physical actions in coordination with humans. Many AI
techniques are invited, from visual processing to symbolic reasoning, from task
planning to \emph{theory of mind} building, from reactive control to action
recognition and learning.

We do not tackle all these issues. This article attempts however to organize
them into a coherent challenge for artificial intelligence, and also to make
explicit some of the paths that have been investigated on our robots and that
result in a deliberative, knowledge-oriented, architecture for human-robot
interaction.

The class of interactions we consider are essentially multi-modal situated
dialog and collaborative task achievement: a robot which is expected to achieve
interactive object manipulation, fetch and carry tasks and similar tasks in a
domestic environment, with human partners (Figure~\ref{fig:hri-dec}). Two
agents (a human and a robot) share a common space and exchange information
through multiple modalities. The robot is expected to explicitly take into
account the human and to manage its interactions with him.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=0.6\columnwidth]{grounding_robot.pdf}
\caption{Robot reasoning about HRI and anticipation of human activities.
  Sources of information are multi-modal dialogue, and monitoring of the
  environment and human activity. The robot must adapt on-line its behaviours
  by merging computed plans with reactive control. The robot explicitly reasons
  on the fact that it is (or not) observed by the human. Reasoning and planning
  take place at symbolic as well as geometric level and systematically take
  into account both agents beliefs and activities (from the point of view of
  the robot).}

\label{fig:hri-dec}
\end{figure}

@@

The remaining of the article discusses the robotic architecture we have built to
tackle the AI challenges of human-robot interaction, and how its relates to
other approaches. We propose to organize this discussion in four sections.

The next section presents the overall architecture and the knowledge model we
have developed for our robots. Section~\ref{sec:impl} gives some insights on
each of the main components of the architecture, and try to highlight their
significance for artificial intelligence. Section~\ref{sec:expe} presents two
experiments that illustrate in a concrete way what can be currently achieved
with our robots.  Section~\ref{sect|conclusion} finally summarizes our main
contributions and relate them to other recent researches in the field.

@@

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

@@

This activity is however not the backbone of our architecture. Other
activities are conducted in parallel, without being explicitly considered as
desires: assessment of the situation and the environment, dialogue (including
performative dialogue that can possibly change the internal state of the robot,
but does not lead directly to the creation of desires, like statement assertion or
question answering), various background monitoring and recognition tasks, etc.

@@

\begin{figure}
        \centering
        \includegraphics[width=0.8\columnwidth]{spark-oro.png}
    \caption{Functional overview of knowledge base ({\sc Oro} server, top part)
    and the geometric situation assessment module (\emph{{\sc Spark}}, bottom
    part). {\sc Spark} computes symbolic relationships between objects and
    agents, and feeds them to the knowledge base.}

@@

   \caption{ The scenario involves a several videotapes that have to be
       manipulated. Other objects are used as supports and containers.  After
       identification and localisation of the set of objects and acquisition of
       the position and posture of the human partner, the robot is able to compute
       that two tapes only are reachable by itself: the \concept{LOTR\_TAPE}
       (black in the 3D model) and \concept{GREY\_TAPE}. The third tape
       \concept{WALLE\_TAPE} (white in the 3D model) and the trashbin
       \concept{PINK\_TRASHBIN} are only reachable by the human. The
       corresponding symbolic facts are listed in Table~\ref{table|beliefsfig7}.
   }%

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\columnwidth]{second_plan.pdf}
  \caption{A plan with the wasted time social rule. {\tt PUTRV} stands for {\it Put it so it is both Reachable and Visible}.}
  \label{plan_hatp2}
\end{figure}

@@

\begin{figure*}[thpb]
  \centering
  \includegraphics[width=1.0\textwidth]{third_plan.pdf} \\
  \caption {A plan produced by HATP to execute the high-level order ``clean the table''. Two streams of actions are generated: for the human (top) and for the robot(bottom). Synchronization points ensure the coordination.}
  \label{plan-etat2}
\end{figure*}

@@

Numerous approaches exist, like amodal \emph{proxies}~\cite{Jacobsson2008},
grounded amodal representations~\cite{Mavridis2006}, semantic maps
(\cite{Nuechter2008, Galindo2008,Blodow2011}) or
affordance-based planning and object classification~\cite{Lorken2008,
Varadarajan2011}.

@@

\paragraph{Planning for cooperation} HRI offers new challenges to planning
by adding a complementary constraint: the synthesis of plans that are not only
feasible but legible (in terms of their intention) and compliant with human
preferences. This constraint is often ill-defined, dynamic, context-dependent.

Indeed, we posit that human-robot collaborative task achievement calls for more
subtlety for a robot than autonomous independent task achievement. HATP, by
co-planning for the robot and the human, and modeling social rules as planning
parameters, is a first step in this direction.

@@

We are planning to extend the scheme by taking into accoount situation where
divergent belief may appear between the human and the robot. Preliminary
results have been pusblished in~\cite{warnier2012when}.

There is also extensive work to be done in order to refine the notion of "good
shared plan" and "good/acceptable robt behaviour" in this context. There are
obviously large avenues for learning and adaptation in this context.
